---
output:
  html_document: default
  pdf_document: default
---
##Predictions

In thinking about predictions, there are a few things to point out:  

+ Voting rates in midterm elections in these counties has been decreasing over the past 15 years
    + Turnout in 2014 was far lower than in 2010, 2006, and 2002
+ Roughly half of the people who turn out for general elections vote in midterms
+ Redistricting has made my predictions very approximate, as I incorporate data from House and Senate midterms of the same year  

Let's make some predictions:  

+ Taking half of the electorate in the 2016 presidential election, we get **38,350 voters** expected in 2018.
    + 2014 had 41,484 voters, so this would follow the decreasing participation rate in Midterms
+ If we take the average of Midterm voters from 2002 to 2014 in the 5 counties, we get **48,500 voters.**

So these are some nice wide bounds, but we want a more accurate estimate. 

We will use two prediction methods and average them together for a decent vote estimate. First, we'll take the turnout rate of the 2014 election and find the expected turnout of 2020, e.g. *(turnout projection)*. Then, we'll take a linear model of the past 4 elections (2002,2006,2010,2012), and calculate a quick linear model for 2018 based on this *(linear projection)*. We can use these together to get a nice range for our estimate. 

```{r,echo=F}
cty_dat <- read.table("countydata.csv",sep = "\t",header = T)
cty_dat[21:25,]$year <- 2018; cty_dat[21:25,]$county <- cty_dat[1:5,]$county
for(i in 1:5){
  sub <- cty_dat[which(cty_dat$county == cty_dat$county[i]),]
  a <- lm(sub$votes[1:4] ~ sub$year[1:4])
  cty_dat[(20+i),]$votes <- a$coefficients[1] + 2018*a$coefficients[2]
}

library(ggplot2)
ggplot(data=cty_dat) + geom_point(aes(year,votes,color=county)) + geom_line(aes(year,votes,color=county)) + scale_y_continuous(limits=c(0,18000),name = "Number of Votes") + scale_x_continuous(breaks=c(2002,2006,2010,2014,2018),name="Election Year",labels=c(02,06,10,14,18))+facet_wrap(~county)
```


```{r}
fteen <- read.csv("popn_14.csv")
```

**Cheatham**:

+ Turnout Projection: 8825  
+ Linear Projection: 8347  
+ *Average*: 8586  
  
**Dickson**:

+ Turnout Projection: 10812
+ Linear Projection: 8347
+ *Average*: 8586
  
**Hickman**:

+ Turnout Projection: 4413
+ Linear Projection: 3743
+ *Average*: 4078
  
**Humphreys**:

+ Turnout Projection: 4152
+ Linear Projection: 3568
+ *Average*: 3860
  
**Robertson**:

+ Turnout Projection: 14654
+ Linear Projection: 14645
+ *Average*: 14649
  
**TOTAL**:

+ Turnout Projection: 42855
+ Linear Projection: 39921
+ *Average*: 41388

So, let's approximate between **40,000 - 41,000**--that's the vote count we expect.
