---
output:
  html_document: default
  pdf_document: default
---

##Predictions

In thinking about predictions, there are a few things to point out:  

+ Voting rates in midterm elections in these counties has been decreasing over the past 15 years
    + Turnout in 2014 was far lower than in 2010, 2006, and 2002
+ Roughly half of the people who turn out for general elections vote in midterms
+ Redistricting has made my predictions very approximate, as I incorporate data from House and Senate midterms of the same year  

Let's make some predictions:  

+ Taking half of the electorate in the 2016 presidential election, we get **38,350 voters** expected in 2018.
    + 2014 had 41,484 voters, so this would follow the decreasing participation rate in Midterms
+ If we take the average of Midterm voters from 2002 to 2014 in the 5 counties, we get **48,500 voters.**

So these are some nice wide bounds, but we want a more accurate estimate. 

We will use two prediction methods and average them together for a decent vote estimate. First, we'll take the turnout rate of the 2014 election and find the expected turnout of 2020, e.g. *(turnout projection)*. Then, we'll take a linear model of the past 4 elections (2002,2006,2010,2012), and calculate a quick linear model for 2018 based on this *(linear projection)*. We can use these together to get a nice range for our estimate. 

##Plots of Votes (including 2018 Projected), by County

---

<center>
```{r,echo=F}
cty_dat <- read.table("data/countydata.csv",sep = "\t",header = T,stringsAsFactors = F)
cty_dat[21:25,]$year <- 2018; cty_dat[21:25,]$county <- cty_dat[1:5,]$county

#fill in linear projection
for(i in 1:5){
  sub <- cty_dat[which(cty_dat$county == cty_dat$county[i]),]
  a <- lm(sub$votes[1:4] ~ sub$year[1:4])
  cty_dat[(20+i),]$votes <- a$coefficients[1] + 2018*a$coefficients[2]
}

#ggplot with changes axes
library(ggplot2)
ggplot(data=cty_dat) + geom_point(aes(year,votes,color=county)) + geom_line(aes(year,votes,color=county)) + scale_y_continuous(limits=c(0,18000),name = "Number of Votes") + scale_x_continuous(breaks=c(2002,2006,2010,2014,2018),name="Election Year",labels=c(02,06,10,14,18))+facet_wrap(~county)
```
</center>

##Table of Projections

---

```{r,echo=F}
fteen <- read.csv("data/popn_14.csv",stringsAsFactors = F)

x <- data.frame(matrix(NA,nrow=5,ncol=3))
colnames(x) <- c("County","LinearProjection","TurnoutProjection")
x$County <- cty_dat[1:5,]$county
x$LinearProjection <- floor(cty_dat[which(cty_dat$year==2018),]$votes)

#calculates % of voting population in 2014 * 2018 expected popn
x$TurnoutProjection <- floor(fteen$popn_18 * (cty_dat[which(cty_dat$year==2014),]$votes/fteen$popn_14))
#write_csv(x,path="x.csv")

#levels(x$County) <- c(levels(x$County), "TOTAL")
#x[6,]$County <- c("TOTAL")
suppressWarnings(library(pander))

totals <- paste("<strong>",c("TOTAL",colSums(x[2:3])),"</strong>")
x <- rbind(x,totals)
#emphasize.strong.rows()

#knitr::kable(rbind(x,".",totals))
#emphasize.strong.cells(totals)
pander(x=x,justify='left')
```

